% !TEX root = ../axiomatic.tex

\section{Introduction}\label{s:introduction}

In \cite{steenrod1947products}, Steenrod introduced by means of formulas the \textit{\mbox{cup-$i$} products} on the cochains of spaces.
These bilinear maps define the celebrated \textit{Steenrod squares}
\[
\Sq^k \colon \rH^\vee(X; \F) \to \rH^\vee(X; \F)
\]
laying at the heart of stable homotopy theory.

Steenrod's formulas for the \mbox{cup-$i$} products extend the Alexander-Whitney product on cochains.
This non-commutative product $\cup_0$ induces the commutative algebra structure on cohomology, and we can interpret the higher \mbox{cup-$i$} products
\[
\cup_i\ \colon \cochains(X; \F)^{\ot 2} \to \cochains(X; \F)
\]
as coherent homotopies witnessing the derived commutativity of $\cup_0$ at the cochain level.

In later work by Steenrod and others, an indirect argument based on the acyclic carrier theorem is used to establish the existence of \mbox{cup-$i$} products and consequently of Steenrod squares.
This approach became the standard since any set of choices for the \mbox{cup-$i$} products homotopic to Steenrod's original gives rise to the same cohomology operations which, by then, had been axiomatically characterized.
As a consequence, the need to interact with a specific set of choices for the \mbox{cup-$i$} products largely declined.

Interest in actual cochain representatives and operations, like the \mbox{cup-$i$} products, has recently been rekindled by their use in the study of topological phases of matter.
A small sample of the physics literature considering this approach is \cite{gaiotto2016spin, kapustin2017fermionic, meng2018classification, wang2020construction, barkeshli2021classification, tata2021anomalies, tata2021cubical}, and some of the mathematical articles motivated by them are \cite{brumfiel2016pontrjagin, brumfiel2018pontrjagin, medina2020cartan, medina2021adem}.
The importance of effective constructions is also highlighted by their use in topological data analysis, where cup-$i$ products have been used in \cite{medina2022per_st} to extract finer features of real-world data -- the conformation space of the cyclo-octane molecule -- using the open-source project \href{https://github.com/Steenroder/steenroder}{\texttt{steenroder}}.

In the present work we introduce an axiomatic characterization of Steenrod's original cup-$i$ construction up to isomorphism -- not just homotopy.
Loosely expressed, our axioms demand that a cup-$i$ construction be natural with respect to simplicial maps, parameterized by the smallest possible complex, not identically $0$ or reducible to subsimplices, and as free as possible with respect to transpositions.

We now say a few words on the relevance of this result.
Together with extending the axiomatic work on cohomology operations \cite{serre1053eilenberg_maclane, cartan1955iteration, steenrod1962cohomology} to the cochain level, it shows that all formulas in the literature \cite{steenrod1947products, real1996computability,mcclure2003multivariable,berger2004combinatorial,medina2020prop1,medina2023fast_sq} defining a cup-$i$ construction represent the same isomorphism class -- for two of these, no previous proof of the equivalence seems to have existed in print.
%Additionally, we show that the operadic constructions of McClure--Smith \cite{mcclure2003multivariable} and Berger--Fresse \cite{berger2004combinatorial} define a bijection between the set of cup-$i$ constructions satisfying our axioms and that of bases of the arity two part of their operads.
This universality highlights the fundamental nature of Steenrod's \mbox{cup-$i$} construction, and it gives more context for its connections with physics -- as mentioned above -- as well as with higher category theory, where the nerve construction can be deduced from Steenrod's construction \cite{street1987orientals, medina2020globular}.
This universality seems grounded in convex geometry, where a cup-$i$ construction on the chains of a convex polytope can be canonically constructed from a orthonormal basis of its normal fan.
%, with Steenrod's construction appearing when considering simplices \cite{medina2022fib_poly}.
%, and $L$-theory, where it is used to algebraically model presheaves on simplicial complexes \cite{medina2022assembly}.

The cup-$i$ products form part of a larger algebraic structure on cochains encoding the entire homotopy type of spaces with certain finiteness assumptions \cite{mandell2006homotopy_type}.
There exists several explicit constructions of these so-called $E_\infty$-algebra structures \cite{mcclure2003multivariable,berger2004combinatorial,medina2020prop1} on simplicial cochains.
For the reader familiar with operad theory, we will review these extensions and observe that the set of bases of the arity $2$ part of the $E_\infty$-operads of McClure--Smith and Berger--Fresse are both in bijective correspondence, through their respective $E_\infty$-algebra structures on simplicial cochains, with the set of cup-$i$ constructions satisfying our axioms.

We conclude this article with a brief discussion -- from the effective viewpoint -- of secondary operations in mod 2 cohomology and Steenrod operations on mod $p$ cohomology.

%We now say a few words on the relevance of this result.
%First, it continues the axiomatic work on cohomology operations \cite{serre1053eilenberg_maclane, cartan1955iteration, steenrod1962cohomology} to the cochain level.
%Second, it clarifies that all available cup-$i$ constructions represent the same isomorphism class.
%For two of these, no previous proof of the equivalence with Steenrod's existed in print.
%%Additionally, we show that the operadic constructions of McClure--Smith \cite{mcclure2003multivariable} and Berger--Fresse \cite{berger2004combinatorial} define a bijection between the set of cup-$i$ constructions satisfying our axioms and that of bases of the arity two part of their operads.
%Third, it highlights the fundamental nature of Steenrod's \mbox{cup-$i$} construction giving more context for its connections with physics, as documented above, higher category theory, where the nerve construction can be deduced from Steenrod's construction \cite{street1987orientals, medina2020globular}.
%Convex geometry, where it is obtained from a orthonormal basis of the normal fan of simplices \cite{medina2022fib_poly}, and $L$-theory, where it is used to algebraically model presheaves on simplicial complexes \cite{medina2022assembly}.

%The first alternative construction of \mbox{cup-$i$} products is due to Real \cite{real1996computability} and it is based on the Eilenberg--Zilber contraction.
%It was further developed by Gonz\'alez-D\'iaz--Real \cite{gonzalez-diaz1999steenrod} and used as the basis for an algorithm, implemented by Palmieri in the open-source project \href{https://www.sagemath.org/}{\texttt{SAGE}}, for the computation of Steenrod squares.
%We will review this construction in \cref{ss:real} and use our axioms to show it is isomorphic -- in fact equal -- to Steenrod's original.
%
%In \cite{medina2023fast_sq}, new formulas defining a \mbox{cup-$i$} construction were introduced and used to present a faster algorithm for the computation of Steenrod squares for finite simplicial complexes.
%The question of comparing the resulting \mbox{cup-$i$} construction with either Steenrod's or Real's was not addressed.
%We use our axiomatic characterization to show in \cref{ss:original} that our construction agrees with Steenrod's.

%A generalization of the notion of \mbox{cup-$i$} product structure is that of $E_\infty$-algebra.
%Using work by \cite[\subsectionSymbol4.5]{benson1998representations} generalizing Steenrod's formulas, McClure--Smith \cite{mcclure2003multivariable} and Berger--Fresse \cite{berger2004combinatorial} constructed natural $E_\infty$-algebra structures on the normalized cochains of simplicial sets.
%As we describe in \cref{ss:operads}, these structures induce a bijection between the set of bases of the arity 2 part of their operads and the set of \mbox{cup-$i$} constructions satisfying our axioms.

%These operadic constructions are obtained by dualizing $E_\infty$-coalgebra structures that extend the Alexander--Whitney diagonal coproduct.
%In \cite{medina2020prop1}, the normalized chains of standard simplicial sets where equipped with a natural $\Med$-bialgebra structure generated by the Alexander--Whitney coproduct, the augmentation map, and an algebraic version of the join of simplices.
%This structure generalizes that of McClure--Smith and Berger--Fresse.
%In particular, the \mbox{cup-$i$} coproducts axiomatized here can be expressed as explicit compositions of two natural maps: the diagonal and the join.
%
%A (non-degenerate) $E_\infty$-coalgebra structure on the normalized chains of simplicial sets defines operations on their mod $p$ cohomology \cite{steenrod1953cyclic, may1970general}.
%When $p = 2$ these agree with Steenrod squares.
%Analogues of Steenrod's \mbox{cup-$i$} coproducts effectively defining these operations were introduced in \cite{medina2021may_st} and implemented in the open-source project \href{https://github.com/ammedmar/comch}{\texttt{ComCH}}.
%These \textit{cup-$(p,i)$ coproducts} are expressible in terms of the diagonal and the join, and we expect to study their moduli in future work, as done here for $p=2$.

\subsection*{Outline}

%\cref{s:preliminaries} reviews the required preliminaries from the theory of simplicial sets.
We present the notion of cup-$i$ construction and our axiomatic characterization in \cref{s:statement}.
We postpone the proof of our main theorem until after \cref{s:reformulation}, where the statement is recasted in terms of a functor naturally isomorphic to that of normalized chains.
\cref{s:proof} is devoted to the proof of our axiomatic characterization.
%The reader is encouraged to start the paper in this section referring back if needed.
%We discuss Steenrod squares in \cref{s:squares}.
%This section, serving to provide context for the subject, is logically independent of the rest of the paper.
%Before presenting any proofs, we recast in \cref{s:reformulation} the main definitions and theorems of \cref{s:statement} in terms of a functor naturally isomorphic to that of normalized chains.
In \cref{s:others} and \cref{s:operads} we show that all formulas in the literature give rise to isomorphic \mbox{cup-$i$} constructions.
We conclude this work in \cref{s:epilogue} discussing related and future work.

