% !TEX root = ../axiomatic.tex

\section{Axioms}\label{s:statement}

In this section we introduce the notion of \mbox{cup-$i$} construction and the axioms that characterize Steenrod's.
We also recall from \cite{medina2023fast_sq} a set of formulas defining -- after the main result of this paper -- Steenrod's cup-$i$ construction.

We denote the (normalized) cochains of a simplicial $X$ by $\cochains(X)$ graded homologically and concentrated in non-positive degrees.
Unless otherwise stated we consider $\Ftwo$ as the field of coefficients.

\subsection{Cup-$i$ constructions}

Let $\sym_2$ be the group with one non-identity element $T$ and let
\[
\begin{tikzcd}[column sep=20pt]
W = \Big(
\Fp[\sym_2]\{e_0\} &
\Fp[\sym_2]\{e_1\} \arrow[l, "\ 1+T"'] &
\arrow[l, "\ 1+T"'] \cdots \Big)
\end{tikzcd}
\]
be the minimal projective resolution of $\F$ by $\F[\sym_2]$-modules.
We notice that the group of automorphisms of $W$ is isomorphic to $\prod_{\N} \sym_2$ with any such $\phi$ defined by a choice for every $i \in \N$ of
\[
\phi(e_i) = e_i
\quad \text{ or } \quad
\phi(e_i) = Te_i.
\]

\begin{definition}
	A \textbf{\mbox{cup-$i$} product structure} on a chain complex $A$ is a chain map
	\[
	W \ot_{\F[\sym_2]} A^{\ot 2} \to A
	\]
	where $T$ acts by transposition on $A^{\ot 2}$ and by right multiplication on $W$.
	We denote the image of $[e_i \ot \alpha \ot \beta]$ by $\alpha \cup_i \beta$.
\end{definition}

Unpacking this structure on $A$ we have the following identity for any $i \in \Z$ and $\alpha, \beta \in A$:
\[
\alpha \cup_{i-1} \beta + \beta \cup_{i-1} \alpha =
\bd (\alpha \cup_{i} \beta) + (\bd\alpha) \cup_{i} \beta + \alpha \cup_{i} (\bd\beta)
\]
with the convention $\alpha \cup_{i} \beta = 0$ for a negative integer $i$.
Setting $i = 0$ we see that $(A, \cup_0)$ is a differential graded algebra.
The identity obtained setting $i = 1$ implies that the algebra structure induced in the homology $HA$ of $A$ is commutative, with the product $\cup_1$ enforcing this relation at the chain level.
This product is itself not commutative, but it is so up to a homotopy defined by $\cup_2$, and so on.
From this chain level structure we deduce the \textit{Steenrod squares} operations $\Sq^k \colon HA \to HA$, defined for $k \in \N$, by the assignment $[\alpha] \mapsto \big[(\alpha \cup_{k-\bars{\alpha}} \alpha)\big]$.

\begin{definition}
	An \textbf{isomorphism} of \mbox{cup-$i$} product structures on $A$ is an automorphism $\phi$ of $W$ making the diagram
	\begin{center}
	\begin{tikzcd}[column sep=5, row sep=15]
	W \displaytensor_{\F[\sym_2]} A \arrow[dr, in=180, out=-90] \arrow[rr, "\phi \, \ot \, \id \, "] & &
	W \displaytensor_{\F[\sym_2]} A \arrow[dl, in=0, out=-90] \\
	& A &
 	\end{tikzcd}
	\end{center}
	commute.
\end{definition}

\begin{definition}
	A \textbf{\mbox{cup-$i$} construction} (more accurately termed \emph{\mbox{cup-$i$} product construction on simplicial cochains}) is a \mbox{cup-$i$} product structure on $\cochains(X)$ for every simplicial set $X$ that is natural with respect to simplicial maps.
	An \textbf{isomorphism} of \mbox{cup-$i$} constructions is an isomorphism of \mbox{cup-$i$} product structures on $\cochains(X)$ for every simplicial set $X$ that is natural with respect to simplicial maps.
\end{definition}

\subsection{Uniqueness}

The first axiom alluded to in the introduction, naturality, has been explicitly included in our definition of \mbox{cup-$i$} construction; whereas the second, minimality, is manifested in the use of $W$ instead of an arbitrary projective resolution of $\F$.

\begin{definition}\label{d:properties}
	A \mbox{cup-$i$} construction is \textbf{non-degenerate} if for any simplex $x$
	\[
	\boxed{x \cup_0 x \neq 0}
	\]
	whenever $\bars{x} = 0$.
	It is \textbf{irreducible} if for any proper face $y$ of $x$
	\[
	\boxed{\Big( y^{(1)} \cup_{i} y^{(2)} \Big)(x) = 0}
	\]
	for any two faces $y^{(1)}$ and $y^{(2)}$ of $y$.
	It is \textbf{free} if for any two simplices $x$ and $y$
	\[
	\boxed{x \cup_{i} y = y \cup_{i} x} \
	\Longrightarrow \
	\boxed{x \cup_{i} y = 0}
	\]
	whenever $\bars{x} \neq i$ or $\bars{y} \neq i$.
\end{definition}

We can now state our main result.

\begin{theorem}\label{t:main}
	There is up to isomorphism only one non-degenerate, irreducible and free \mbox{cup-$i$} construction.
\end{theorem}

We will use this result to prove in \cref{s:others,s:operads} that all \mbox{cup-$i$} constructions in the literature are isomorphic to Steenrod's original \cite{steenrod1947products}.

\subsection{Existence}

To prove our main result we will use the formulas defining one such \mbox{cup-$i$} construction introduced in \cite{medina2023fast_sq}, and which we now review.

\begin{notation*}
	We will use the traditional notation for face maps, and for any $n$-simplex $x$ and set $U = \{u_1 < \dots < u_r\} \subseteq \big\{ 0, \dots, n \big\}$ we write $d_U(x)$ for $d_{u_1}\! \dotsm d_{u_r}(x)$ with $d_{\emptyset}(x) = x$.
\end{notation*}

\begin{definition}[\cite{medina2023fast_sq}] \label{d:my cup-i}
	Let $X$ be a simplicial set, $x \in X_n$ and $\alpha, \beta \in \cochains(X)$.
	\begin{equation}\label{e:new formulas}
	(\alpha \cup_i \beta)(x) =
	\begin{cases}
	(\alpha \ot \beta) \sum d_{U^0}(x) \ot d_{U^1}(x) &
	\text{if } i \in \{0, \dots, n\}, \\
	\hfil 0 &
	\text{otherwise},
	\end{cases}
	\end{equation}
	where the sum is taken over all $U = \{u_1 < \cdots < u_{n-i}\} \subseteq \{0, \dots, n\}$ and
	\begin{equation*}
	U^0 = \{u_j \in U \mid u_j \equiv j \text{ mod } 2\}, \qquad
	U^1 = \{u_j \in U \mid u_j \not\equiv j \text{ mod } 2\}.
	\end{equation*}
\end{definition}

\begin{example}
	For $i = 0$, \cref{e:new formulas} gives
	\begin{equation*}
	(\alpha \cup_0 \beta)(x) =
	\sum_{j=0}^n \alpha \big(d_{j+1} \cdots d_{n}(x)\big) \cdot \beta \big(d_{0} \cdots d_{j-1}(x)\big),
	\end{equation*}
	the well-known \textbf{Alexander--Whitney product}.
\end{example}

The verification that the formulas in \cref{e:new formulas} define a \mbox{cup-$i$} construction is given in \cite{medina2023fast_sq}.
We will refer to it as the \textbf{canonical \mbox{cup-$i$} construction}.
Furthermore, we will show that it is non-degenerate, irreducible, and free (\cref{t:existence}) and, using our main theorem, that it agrees with Steenrod's original cup-$i$ construction (\cref{t:steenrod cup-i}).